---
title: "analysis"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(bayesplot)
library(boot)
library(brms)
library(here)
library(latex2exp)
library(scales)
library(tidybayes)
library(tidyverse)
source("utils.R")

theme_set(theme_minimal(base_size=11) + 
            theme(legend.position = "top", panel.spacing = unit(2, "lines")))
target_dir = here("results")

save_plots <- FALSE
plot_dir = file.path(target_dir, "figs")
if(!file.exists(plot_dir)) dir.create(plot_dir, recursive = TRUE)
```

# Load cleaned experimental data

Data was cleaned (see script R/clean_data.R) according to predefined exclusion criteria

```{r, message = FALSE, warning = FALSE}
######## select data #######
fn_data = "data_cleaned.csv"
fn_model = "statistical-ordinal-brms-model.rds"

# fn_data = "cleaned_data_not_always_both.csv"
# fn_model = "statistical-ordinal-brms-model-not-always-two-selected.rds"

data = read_csv(here("results", fn_data)) %>% group_by(prolific_id)
ordinal_model <- readRDS(here("results", fn_model))
```

```{r}
message(paste("model:", fn_model))
message(paste("data:", fn_data))

data.test = data %>% filter(str_detect(block, "test"))
data.practice = data %>% filter(str_detect(block, "practice"))

data.critical <- data %>% filter(type == "critical") %>% 
  dplyr::select(submission_id, prolific_id, id, type, QUD, response, 
                selected_pic, pic2, pic3, RT) %>% 
  mutate(pic2=str_replace(pic2, ".png", ""), pic2=str_replace(pic2, "_horiz", ""),
         pic3 = str_replace(pic3, ".png", ""), pic3=str_replace(pic3, "_horiz", "")) %>% 
  mutate(picPair=paste(pic2, pic3, sep=" & "),
         pic_condition = 
           case_when(picPair == "if2_unn & if2_unu" ~ "D. withDistractor-external", 
                     picPair == "if1_un & if2_unu" ~ "B. w/oDistractor-external",
                     picPair == "if1_un & if1_uu" ~ "A. w/oDistractor-internal",
                     picPair == "if2_unn & if1_uu" ~ "C. withDistractor-internal")) %>% 
  dplyr::select(-pic2, -pic3) %>%
  mutate(stimulus = substr(pic_condition, 1, 1),
         pic_condition = substr(pic_condition, 4, nchar(pic_condition)), 
         response = case_when(response == "exhaustive_non-exhaustive" ~ "both",
                              T ~ response),
         QUD = case_when(QUD == "willq" ~ "will-q",
                         QUD == "ifp" ~ "if-p", 
                         T ~ QUD))
```

## Participants' responses to questions shown in the end

(Multiple selections possible)
During the experiment, Ann's question was ...

- always the same (**same**)
- sometimes about what happened if one block fell (**ifp**)
- sometimes about the yellow block (**yellow**)
- sometimes about whether one block would fall (**willq**)
- sometimes about several blocks at a time (**several**)
- I only read Bob's answer (**ignored**)

```{r}
data %>% dplyr::select(prolific_id, check_ann) %>%
  distinct() %>% group_by(check_ann) %>% dplyr::count() %>% arrange(desc(n))
```

Most of the time I selected a **single picture** ...

- even though another picture wasn't unlikely either
- only when I was very confident about my decision

```{r}
data %>% dplyr::select(prolific_id, check_behavior) %>%
  distinct() %>% group_by(check_behavior) %>% dplyr::count()
```

## Trials where a pragmatic and a literal selection is possible

For instance in one filler trial, Bob says: "The green block will NOT fall" and
in one scene, all other blocks fall (referred to as 'pragmatic'), in another
scene, some other block do not fall either (referred to as 'literal').

```{r}
trials_lit_prag = c("trial7", "trial8", "trial19", "trial20", "trial21", "trial24")
data.lit_prag <- data %>% filter(id %in% trials_lit_prag) %>% 
  dplyr::select(prolific_id, response, type, id, response) %>% group_by(id)

# by trials
data.lit_prag %>% 
  ggplot(aes(x=id, fill=response)) + 
  geom_bar(stat="count", position=position_dodge()) +
  scale_y_continuous(breaks = pretty_breaks()) +
  facet_wrap(~type, scales = "free_x") +
  guides(fill = guide_legend(ncol = 3))
```

## "Pragmatic" scores

Scores for participants how pragmatic their responses were. 
1: pragmatic picture selected (pragmatic)
0.5: pragmatic + literal picture selected (literal2)
0: literal picture selected (literal1)

Max score:6
Min score:0

```{r}
data.pragmatic_score = data.lit_prag %>% 
  mutate(score = case_when(response == "pragmatic" ~ 1,
                           response == "literal" ~ 0, 
                           T ~ 0.5)) %>% 
  group_by(prolific_id) %>% summarize(score = sum(score), .groups = "drop") %>%
  mutate(relative_score = round(score / 6, 2))

data.pragmatic_score %>% 
  ggplot(aes(x=score)) + geom_bar() + 
  labs(x = "pragmatic score", y = "nb participants") +
  scale_x_continuous(breaks = pretty_breaks())

data.pragmatic_score %>% summarize(abs = mean(score), rel = mean(relative_score))
```

## sanity check: control scene

(should be an empty dataframe) 
Note that in trial10 (filler) two scenes called 'contrast', just pic1 is 
the control scene participants should not select (in other 'contrast'-scene, 
utterance is indeed literally true)

```{r}
data.test %>% 
  dplyr::select(prolific_id, id, type, response, selected_pic) %>% 
  filter(str_detect(response, "contrast") & str_detect(selected_pic, "pic1"))
```

## Practice trials

```{r}
data.practice %>%
  ggplot(aes(x=id, fill=response)) + 
  geom_bar(stat="count", position=position_dodge()) +
  scale_y_continuous(breaks = pretty_breaks()) +
  guides(fill = guide_legend(nrow = 4))
```

## Trials where we'd expect selection of exhaustive + non-exhaustive

Trials, where a selection of one is risky since no clue for neither of both
Trial11 is a practice trial, trial13 a filler trial.

```{r}
risk_trials = c("trial11", "trial13")
data.risk <- data %>% filter(id %in% risk_trials) %>% 
  dplyr::select(prolific_id, response, id, type)

data.risk %>% ggplot(aes(x=id, fill=response)) + 
  geom_bar(stat="count", position=position_dodge()) +
  scale_y_continuous(breaks= pretty_breaks()) +
  guides(fill = guide_legend(nrow=2))

```

## Meta information about participants

```{r, meta, echo=FALSE}
data.meta <- data %>% 
  distinct_at(vars(c(submission_id)), .keep_all = T) %>%
  mutate(education=as_factor(education), gender=as_factor(gender))

data.meta$gender %>% summary()
data.meta$education %>% summary()
print("age:")
data.meta$age %>% summary()
print("timeSpent:")
data.meta$timeSpent %>% summary()

data.meta %>% ggplot(aes(x="time", y=timeSpent)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter() +
  labs(x="")
```

## Participants' game scores from test phase

Feedback given in practice trials, in other trials particapants do not see their scores. 

```{r}
data.test_scores = data %>% filter(type != "training") %>% 
  dplyr::select(prolific_id, score, type, id) %>%
  group_by(type, prolific_id) %>%
  summarize(score = sum(score), .groups = "drop_last")

xticks = data.test_scores$score %>% unique %>% sort()
data.test_scores %>% mutate(score = score / 100) %>% 
  ggplot(aes(x=score, fill=type)) + 
  geom_bar(stat="count") + labs(y = "nb participants") + 
  scale_x_continuous(breaks=xticks/100) +
  facet_wrap(~type, scales = "free") +
  theme(axis.text.x = element_text(size = 07))
```

## Training data

Sanity check: does ratio of correct answers increase with increasing trial numbers?

```{r}
data.train = data %>% filter(type=="training") %>%
  mutate(correct=case_when(expected == response ~ T, T ~ F))

# ratio correct by participants
data.train %>% dplyr::select(prolific_id, id, correct) %>% 
  summarize(nb_trials = n(), nb_correct = sum(correct)) %>% 
  mutate(ratio_correct = nb_correct / nb_trials) %>% 
  ggplot(aes(x=ratio_correct)) + geom_bar(stat="count") + 
  labs(y = "nb participants")

# ratio correct per trial number (different order per participant)
train.correct.nb = data.train %>% group_by(trial_number) %>% 
  summarize(mean_correct = mean(correct))
train.correct.nb %>% 
  ggplot(aes(x=trial_number, y=mean_correct)) + 
  geom_bar(stat="identity") + 
  scale_x_continuous(breaks = scales::pretty_breaks(8)) +
  geom_hline(aes(yintercept = 0.5))

# ratio correct per trial id
train.correct.id = data.train %>% group_by(id) %>% 
  summarize(mean_correct = mean(correct))
train.correct.id %>% 
  ggplot(aes(x=id, y=mean_correct)) + 
  geom_bar(stat="identity") +
  theme(axis.text.x=element_text(angle=90)) + 
  geom_hline(aes(yintercept = 0.5))
```

# Critical test trials

Plot critical trials with 95%-bootstrap confidence intervals

Get bootstrap intervals

```{r}
df.critical <- data.critical %>% 
  group_by(pic_condition, stimulus, QUD) %>% 
  dplyr::select(submission_id, pic_condition, stimulus, QUD, response)

df.crit.exh = df.critical %>% 
  mutate(response = case_when(response == "exhaustive" ~ 1, T ~ 0))
df.crit.nonExh = df.critical %>% 
  mutate(response = case_when(response == "non-exhaustive" ~ 1, T ~ 0))
df.crit.both = df.critical %>% 
  mutate(response = case_when(response == "both" ~ 1, T ~ 0))
df.crit.other = df.critical %>% 
  mutate(response = case_when(!response %in% c("exhaustive", "non-exhaustive", 
                                               "both") ~ 1, T ~ 0))
bootstrap_samples = bind_rows(
  get_bootstrap_statistics(df.crit.both, "both"),
  get_bootstrap_statistics(df.crit.exh, "exhaustive"),
  get_bootstrap_statistics(df.crit.nonExh, "non-exhaustive"),
  get_bootstrap_statistics(df.crit.other, "other selection")
) %>% group_by(pic_condition, stimulus, QUD, response)

CIs <- get_bootstrap_cis(bootstrap_samples)
CIs.across_BD = CIs %>% filter(stimulus == "B" | stimulus == "D") %>% 
  group_by(QUD, response) %>% 
  summarize(ci.low = mean(ci.low), ci.up = mean(ci.up), .groups = "drop_last")

QUDs = list(neutral = "Which blocks do you think will fall?",
            `if-p` = "What happens if the 'antecedent-block' falls?",
            `will-q` = "Will the 'consequent-block' fall?")
```

## Plotted across all different picture combinations (just stimulus B + D)
to compare QUD:will-q vs. QUD:if-p

```{r}
df.critical.BD = data.critical %>% filter(stimulus == "B" | stimulus == "D") %>% 
  mutate(response = factor(response, levels = c("non-exhaustive", "both",
                                                "exhaustive"))) 
df.critical.BD %>% 
  ggplot(aes(fill=response, y=QUD)) + 
  geom_bar(position=position_dodge(preserve = 'single')) +
  scale_y_discrete(breaks=names(QUDs)) +#, labels = QUDs) + 
  labs(y="Ann's question", title="across stimuli B + D") +
  guides(fill = guide_legend(ncol = 3))

# same plot with exhaustive-picture selection ratios instead of absolute numbers
data.critical.rate.BD = df.critical.BD %>% group_by(QUD, response) %>%
  summarize(n=n(), .groups = "drop_last") %>% mutate(ratio = n/sum(n))
data.critical.rate_with_cis.BD = 
  left_join(data.critical.rate.BD, CIs.across_BD, by=c("QUD", "response")) %>%
  mutate(response = factor(response, levels = c("non-exhaustive", "both",
                                              "exhaustive")), QUD = as.factor(QUD))

data.critical.rate_with_cis.BD %>% 
  ggplot(aes(x=ratio, y=QUD, fill=response, group = response)) + 
  geom_bar(stat="identity", position = position_dodge()) + 
  geom_errorbar(aes(xmin=ci.low, xmax=ci.up), 
                position = position_dodge()) +
  geom_text(aes(label=round(ratio, 2)), hjust=-0.25, vjust=-0.45, color = 'blue',
            position = position_dodge(width = .9)) +
  scale_y_discrete(breaks=names(QUDs)) + #, labels = QUDs) + 
  labs(y="Ann's question", x="selection rate", title="across stimuli B + D") +
  guides(fill = guide_legend(ncol = 2))

# again as point plot, not bar plot
w_dodge = 0.25
p = data.critical.rate_with_cis.BD %>% group_by(QUD, response) %>% 
  ggplot(aes(y=ratio, x=QUD, color=response, group=response)) + 
  geom_point(position = position_dodge(width=w_dodge)) +
  geom_line(position = position_dodge(width = w_dodge)) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.up), width=0.1, 
                position = position_dodge(width=w_dodge)) +
  labs(y="selection rate", x = "QUD", title="across stimuli B + D")
p
```

## Plotted separately for each picture combination

```{r}
# same plot with exhaustive-picture selection ratios instead of absolute numbers
data.critical.rate.by_stimulus = data.critical %>%
    # filter(!prolific_id %in% ids.critical_always_exhaustive & 
    #        !prolific_id %in% ids.selected_always_two) %>% 
  group_by(pic_condition, stimulus, QUD, response) %>%
  summarize(n=n(), .groups = "drop_last") %>% mutate(ratio = n/sum(n))

data.critical.rate.by_stimulus_with_cis =
  left_join(data.critical.rate.by_stimulus, CIs, 
            by = c("pic_condition", "stimulus", "QUD", "response")) %>%
  mutate(response = factor(response, 
                           levels = c("non-exhaustive", "both", "exhaustive")),
         QUD = as.factor(QUD))

p = data.critical.rate.by_stimulus_with_cis %>% 
  ggplot(aes(x=ratio, y=QUD, fill=response, group=response)) + 
  geom_bar(stat="identity", position = position_dodge()) + 
  geom_errorbar(aes(xmin=ci.low, xmax=ci.up), position = position_dodge()) +
  geom_text(aes(label=round(ratio, 2)), hjust=-0.25, vjust=-0.3,
            position = position_dodge(width = .9), size = 2.5, color='blue') +
  facet_wrap(~stimulus) +
  labs(x="selection rate", y = "QUD")
p
if(save_plots) ggsave(here(plot_dir, "selection-rates-barplot.pdf"), p, 
                      width=7, height=4)

w_dodge = 0.25
p = data.critical.rate.by_stimulus_with_cis %>% 
  # filter(stimulus %in% c("B", "D")) %>% 
  ggplot(aes(y=ratio, x=QUD, color=response, group=response)) + 
  geom_point(position = position_dodge(width=w_dodge)) +
  geom_line(position = position_dodge(width = w_dodge)) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.up), width=0.1, 
                position = position_dodge(width=w_dodge)) +
  facet_wrap(~stimulus, scales = "free") + 
  labs(y="selection rate", x = "QUD")
p
data.critical.rate.by_stimulus_with_cis %>% 
  filter(stimulus %in% c("B", "D"))
if(save_plots) ggsave(here(plot_dir, "selection-rates-pointplot.pdf"), 
                      p, width=5, height=4)


point2 <- data.critical.rate.by_stimulus_with_cis %>% 
  # filter(stimulus %in% c("B", "D")) %>% 
  ggplot(aes(y=ratio, x=response, color=QUD, group=QUD)) + 
  geom_point(position = position_dodge(width=w_dodge)) +
  geom_line(position = position_dodge(width = w_dodge)) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.up), width=0.1, 
                position = position_dodge(width=w_dodge)) +
  theme(axis.text = element_text(size = 26), 
        axis.title = element_text(size = 26), 
        legend.text = element_text(size = 26), 
        legend.title = element_text(size = 26)) +
  facet_wrap(~stimulus, scales = "free") + 
  labs(y="selection rate", x = "QUD")
point2
if(save_plots) ggsave(here(plot_dir, "selection-rates-pointplot2.pdf"), point2, 
                      width=16, height=8)


point3 <- data.critical.rate.by_stimulus_with_cis %>% 
  filter(stimulus %in% c("B", "D")) %>% 
  ggplot(aes(y=ratio, x=QUD, color=stimulus, group=stimulus)) + 
  geom_point(position = position_dodge(width=w_dodge)) +
  geom_line(position = position_dodge(width = w_dodge)) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.up), width=0.1, 
                position = position_dodge(width=w_dodge)) +
  facet_wrap(~response) #+ 
  # labs(y="selection rate", x = "QUD")
point3
if(save_plots) ggsave(here(plot_dir, "selection-rates-pointplot3.pdf"), point3, 
                      width=6, height=4)

```


##  Responses for QUD:will-q and QUD:if-p plotted together

```{r}
get_responses_all_QUDs = function(dat){
  df = dat %>% group_by(submission_id, pic_condition, stimulus) %>% 
    dplyr::select(submission_id, prolific_id, pic_condition, stimulus, QUD, response)

  df.wide = df %>% group_by(pic_condition, stimulus, submission_id) %>% 
    mutate(response = str_replace(response, "non-exhaustive", "NE"), 
           response = str_replace(response, "exhaustive", "E")) %>% 
    pivot_wider(names_from = QUD, values_from = response, values_fill = "-") %>% 
    unite("ifp.willq", `if-p`, `will-q`, sep="/") %>% 
    separate(pic_condition, into = c("exh", "nonExh"), sep = "-") %>%
    rename(response = ifp.willq)
  return(df.wide)
}

plot_responses_all_QUDs = function(dat) {
  df.wide <- get_responses_all_QUDs(dat)
  df.wide.count = df.wide %>% group_by(stimulus, exh, nonExh, response) %>% 
    dplyr::count() %>% group_by(stimulus) %>% mutate(N=sum(n), ratio=n/N)
  
  xlab = "selected situations for both QUDs: if-p/will-q"
  p <- df.wide.count %>% ggplot(aes(y=response, x=ratio , fill=response)) + 
    geom_bar(stat="identity", position=position_dodge()) +
    geom_label(aes(label=n)) +
    scale_fill_brewer(palette='Set3', name="response ifp_willq") +
    facet_wrap(~stimulus, nrow = 2, scales = "free") +
    theme(legend.position = "none") +
    labs(y = "ratio participants", x=xlab)
  return(p)
}
p <- plot_responses_all_QUDs(data.critical)
p
ggsave(file.path(plot_dir, "empirical-responses.pdf"), p, width=7, height=4)
```


## Scores concerning the nb. of times participants give different responses depending on QUD

Score participants: for the two critical stimuli that occur with both QUDs (B +
D), participants receive 1 point if they differentiate the QUDs ifp_willq, i.e.
when they make different selections for the two QUDs. A score of 0 then means that the
respective participant never made a difference between the two QUDs and a score
of two that participant always made a difference between them.

```{r}
responses.all_quds = get_responses_all_QUDs(data.critical) %>% 
  separate("response", into = c("ifp", "willq"), sep="/") %>% 
  mutate(score = case_when(ifp == willq ~ 0, T ~ 1)) %>%
  filter(willq != "-") # some stimuli are not included with QUD=will-q

df.qud_scores = responses.all_quds %>% group_by(submission_id, prolific_id) %>%
  summarize(score.total=sum(score), .groups = "drop_last")
df.qud_scores.summary = df.qud_scores %>% group_by(score.total) %>% 
  dplyr::count() %>% ungroup() %>% mutate(N=sum(n)) %>% 
  mutate(ratio = n/N)

df.qud_scores.summary %>% ggplot(aes(x=score.total, y=ratio)) + 
  geom_bar(stat="identity") + geom_text(aes(label = round(ratio, 2))) +
  labs(x="nb times participants gave different answers for different QUDs", 
       y = "nb participants")

# never differentiate QUDs
ids.score0 <- df.qud_scores %>% filter(score.total == 0) %>% pull(submission_id)
plot_responses_all_QUDs(data.critical %>% filter(submission_id %in% ids.score0)) +
  ggtitle("Participants who always ignored QUD")

# sometimes differentiate QUDs
ids.score1_or_2 <- df.qud_scores %>% filter(score.total >= 1) %>% pull(submission_id)
plot_responses_all_QUDs(data.critical %>% filter(submission_id %in% ids.score1_or_2)) +
  ggtitle("Participants who sometimes ignored QUD")

# always differentiate QUDs
ids.score2 <- df.qud_scores %>% filter(score.total == 2) %>% pull(submission_id)
plot_responses_all_QUDs(data.critical %>% filter(submission_id %in% ids.score2)) +
  ggtitle("Participants who never ignored QUD")
```

Those who never take QUD into account, do they switch between exh-exh and 
nonExh-nonExh or give the same (identical) answer for both, stimulus B and 
stimulus D?

```{r}
df.ignored_qud = responses.all_quds %>% filter(submission_id %in% ids.score0) %>% 
  group_by(submission_id, ifp) %>% 
  dplyr::count(name="nb_stimuli") %>%
  arrange(submission_id) %>% rename(selection = ifp) %>% ungroup() %>%
  mutate(switch_answer = case_when(nb_stimuli == 2 ~ FALSE, TRUE ~ TRUE))

df.ignored_qud %>% distinct_at(vars(c(submission_id)), .keep_all = T) %>% 
  group_by(switch_answer) %>% 
  dplyr::count() %>% ungroup() %>% mutate(N=sum(n), ratio = n/N)
```

## Plot empirical selection rates

```{r}
df <- data.critical.rate.by_stimulus_with_cis %>% 
  separate(col = pic_condition, into = c("exh_situation", "nonExh_situation"), sep="-") %>%
  mutate(exh_situation = factor(exh_situation, levels = c("withDistractor", "w/oDistractor")),
         response = str_replace(response, "non-exhaustive", "nonExh"),
                    response = str_replace(response, "exhaustive", "exh"))

w_dodge = 0.25
p1 <- df %>%
  ggplot(aes(x=exh_situation, y=ratio, color = response, shape = stimulus)) +  
             # group=interaction(QUD, nonExh_situation))) +
  geom_point(position = position_dodge(width=w_dodge)) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.up), width=0.1, 
                position = position_dodge(width=w_dodge)) +
  # geom_line(position = position_dodge(width=w_dodge)) +
  # labs(y = "selection rates", x = "exhaustive situation") +
  scale_color_brewer(name = "selection", palette = "Set2") +
  guides(color = guide_legend(title.position="top"), 
         shape = guide_legend(title = "stimulus", title.position="top")) +
  facet_grid(QUD ~ nonExh_situation, scales = "free_x", labeller = label_both)
p1
ggsave(file.path(plot_dir, "empirical-responses-rates-fig1.pdf"), p1, 
       width=7, height=4)

p2 <- df %>% 
  ggplot(aes(x=stimulus, y=ratio, color = response)) +
             #group=interaction(QUD, nonExh_situation))) +
  geom_point(position = position_dodge(width=w_dodge)) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.up), width=0.1, 
                position = position_dodge(width=w_dodge)) +
  #geom_line(position = position_dodge(width=w_dodge)) +
  #labs(y = "selection rates", x = "exhaustive situation") +
  labs(x="stimulus", y="selection rate") +
  scale_color_brewer(name = "selection", palette = "Set2") +
  #guides(color = guide_legend(title.position="top")) +
         #shape = guide_legend(title = "stimulus", title.position="top")) +
  facet_grid(QUD ~ stimulus, scales = "free", labeller = label_both)

p2
ggsave(file.path(plot_dir, "empirical-responses-rates-fig2.pdf"), p2, 
       width=7, height=4)

p3 <- df %>% 
  ggplot(aes(x=QUD, y=ratio, color = response, group = response)) +
             #group=interaction(QUD, nonExh_situation))) +
  geom_point(position = position_dodge(width=w_dodge)) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.up), width=0.1, 
                position = position_dodge(width=w_dodge)) +
  geom_line(position = position_dodge(width=w_dodge)) +
  labs(x="stimulus", y="selection rate") +
  scale_color_brewer(name = "selection", palette = "Set2") +
  facet_grid(~ stimulus, scales = "free", labeller = label_both)

p3
ggsave(file.path(plot_dir, "empirical-responses-rates-fig3.pdf"), p3, 
       width=7, height=4)

```


# Bayesian regression models

## Bayesian ordinal regression model

###  Main analysis 

```{r}
df.ordinal <- ordinal_model$data %>% as_tibble() %>% 
    mutate(response = factor(response, levels=c("NE", "both", "E"), ordered=T),
           response_int = as.integer(response)) %>% 
  mutate(exhaustive = exh, non_exhaustive = nonExh) %>% 
  map_conditions_to_stimuli()

yrep = posterior_predict(ordinal_model)
ppc_bars_grouped(y = df.ordinal$response_int, yrep = yrep, group = df.ordinal$QUD)
ppc_bars_grouped(y = df.ordinal$response_int, yrep = yrep, group = df.ordinal$stimulus)

pp_check(ordinal_model, type = 'hist')
pp_check(ordinal_model, type= 'stat', stat = 'mean', nsamples = 100)

conds_stimuli <- make_conditions(df.ordinal, c("exh", "nonExh"))
effects <- conditional_effects(ordinal_model, c("QUD"), categorical=T, 
                               conditions=conds_stimuli, robust=T, re_formula=NA)
effects 

cond_effects = effects$`QUD:cats__` %>% as_tibble() %>% 
  mutate(response = as.character(cats__),
         response = case_when(response == "NE" ~ "non-exhaustive",
                              response == "E" ~ "exhaustive", 
                              T ~ response)) %>% 
  rename(exhaustive = exh, non_exhaustive = nonExh,
         ratio = `estimate__`, ci.low = `lower__`, 
         ci.up = `upper__`) %>% 
  map_conditions_to_stimuli() %>% 
  dplyr::select(stimulus, QUD, response, ratio, ci.low, ci.up)

#prediction for will-q larger than for ifp in respective stimuli
hypothesis(ordinal_model, "QUDwillMq > 0") # stimulus D
hypothesis(ordinal_model,  "QUDwillMq + exhwoD + QUDwillMq:exhwoD > exhwoD") # B
```

<!-- Plot predictions with observed data -->

```{r}
# df <- bind_rows(data.critical.rate.by_stimulus_with_cis %>% 
#                   ungroup() %>% dplyr::select(-pic_condition, -n) %>% 
#                   add_column(data = "observed"), 
#                 cond_effects %>% add_column(data = "predicted"))
# w_dodge = 0.25
# 
# p = df %>%
#   ggplot(aes(y=ratio, x=QUD, color=data, group=interaction(response, data), 
#              shape = response)) + 
#   geom_point(position = position_dodge(width=w_dodge)) +
#   geom_line(position = position_dodge(width = w_dodge)) +
#   geom_errorbar(aes(ymin=ci.low, ymax=ci.up), width=0.1, 
#                 position = position_dodge(width=w_dodge)) +
#   facet_wrap(~stimulus, scales = "free_x") + 
#   labs(y="selection rate", x = "QUD")
# p
# ggsave(file.path(plot_dir, "observed-with-predictions.pdf"), p, width=7, height=5)
```


To test effect of QUD across stimuli, draw samples from posterior

```{r}
posterior_preds = function(model, qud=NA, exh=NA, non_exh=NA, prob_scale=F) {
  qud = str_replace(qud, "-", "") # to catch if-p/will-q inputs as well
  samples_posterior = model %>% 
    spread_draws(`b_Intercept[1]`, `b_Intercept[2]`, `b_QUDwillMq`, 
                 `b_exhwoD`, `b_nonExhint`, `b_QUDwillMq:exhwoD`,
                 `b_exhwoD:nonExhint`) %>% 
    rename(tau1 = `b_Intercept[1]`, tau2 = `b_Intercept[2]`) %>% 
    mutate(
      eta_willq_withD_ext = `b_QUDwillMq`,
      eta_ifp_withD_int = `b_nonExhint`,
      eta_ifp_woD_ext = `b_exhwoD`,
      
      eta_ifp_woD_int = `b_exhwoD` + `b_nonExhint` + `b_exhwoD:nonExhint`,
      eta_willq_woD_ext = `b_QUDwillMq` + `b_exhwoD` + `b_QUDwillMq:exhwoD`
    )
  
  if(prob_scale) {
    eta = paste("eta", qud, exh, non_exh, sep = "_")
    if(eta == "eta_ifp_withD_ext") {
      # predictions for reference category
      predictions = samples_posterior %>% 
        mutate(NE = pnorm(tau1), 
               both = pnorm(tau2) - pnorm(tau1), 
               E = 1 - pnorm(tau2))
      
    } else {
      predictions = samples_posterior %>%
        mutate(NE = pnorm(tau1 - .data[[eta]]), 
               both = pnorm(tau2 - .data[[eta]]) - pnorm(tau1 - .data[[eta]]),
               E = 1 - pnorm(tau2 - .data[[eta]])) 
    }
    samples_posterior <- predictions %>% 
      dplyr::select(starts_with("."), NE, both, E) %>% 
      add_column(QUD=qud, exhaustive=exh, non_exhaustive=non_exh)
  } 
  return(samples_posterior)
}

# based on coefficients
samples_posterior_coeffs = posterior_preds(ordinal_model)
samples_posterior_coeffs %>%
  summarize(p_main.D = mean(b_QUDwillMq > 0),
            p_main.B = mean(eta_willq_woD_ext > eta_ifp_woD_ext), 
            p_main = mean(c(p_main.B, p_main.D)))

# same but based on predicted probabilities
samples_posterior_probs =
  bind_rows(posterior_preds(ordinal_model, "willq", "withD", "ext", T),
            posterior_preds(ordinal_model, "willq", "woD", "ext", T),
            posterior_preds(ordinal_model, "ifp", "withD", "ext", T),
            posterior_preds(ordinal_model, "ifp", "woD", "ext", T),
            posterior_preds(ordinal_model, "ifp", "withD", "int", T),
            posterior_preds(ordinal_model, "ifp", "woD", "int", T)) %>% 
  map_conditions_to_stimuli() 

df.main_hypothesis = samples_posterior_probs %>% 
  dplyr::select(.draw, QUD, stimulus, E) %>% 
  group_by(QUD, stimulus) %>% 
  pivot_wider(names_from = "QUD", values_from = "E") %>% 
  mutate(hypothesis = willq > ifp) 

df.main_hypothesis %>% summarize(p = mean(hypothesis)) %>% 
  pivot_wider(names_from = "stimulus", values_from = "p") %>% 
  mutate(p_main = mean(c(B, D)))

```
Note: we looked at the selection rate of the exhaustive situation, looking at the selection rate of both (exhaustive + non-exhaustive) yields the same results since the third option (non-exhaustive situation) is rarely selected.

```{r}
df.both = samples_posterior_probs %>% 
  filter(stimulus == "B" | stimulus == "D") %>% 
  dplyr::select(.draw, QUD, stimulus, both) %>% 
  group_by(QUD, stimulus) %>% 
  pivot_wider(names_from = "QUD", values_from = "both") %>% 
  mutate(hypothesis_both = ifp > willq) %>% 
  summarize(hypothesis_both = mean(hypothesis_both))
df.both
```


### Posterior predictive checks

```{r}
yrep <- posterior_predict(ordinal_model, ndraws = 1000)

plot_posterior_check_by_stimulus = function(dat, stimulus_letter){
  df <- dat %>% rowid_to_column() %>% filter(stimulus == !!stimulus_letter)
  
  if(stimulus_letter %in% c("B", "D")) {
    indices.if = df %>% filter(QUD == "if-p") %>% pull(rowid)
    indices.willq = df %>% filter(QUD == "will-q")  %>% pull(rowid)
  
    pp_samples.if =  yrep[,indices.if] %>% as_tibble()
    df.if <- pp_samples.if %>% rowid_to_column() %>% pivot_longer(cols=-rowid) %>% 
      group_by(rowid, value) %>% summarize(n=n(), .groups = "drop_last") %>% 
      mutate(p = n/sum(n)) %>% add_column(QUD = "if-p") %>% rename(response_cat = value)
      
    pp_samples.willq =  yrep[,indices.willq] %>% as_tibble()
    df.willq <- pp_samples.willq %>% rowid_to_column() %>% pivot_longer(cols=-rowid) %>% 
      group_by(rowid, value) %>% summarize(n=n(), .groups = "drop_last") %>% 
      mutate(p = n/sum(n)) %>% add_column(QUD = "will-q")%>% rename(response_cat = value)
    
    df.samples <- bind_rows(df.if, df.willq)
    
  } else {
    df.samples <- yrep[, df %>% pull(rowid)] %>% as_tibble() %>%
      rowid_to_column() %>% pivot_longer(cols=-rowid) %>% 
      group_by(rowid, value) %>% summarize(n=n(), .groups = "drop_last") %>% 
      mutate(p = n/sum(n)) %>% add_column(QUD = "if-p") %>% rename(response_cat = value)
  }
  
  pp_samples <- df.samples %>% 
    mutate(response = case_when(response_cat == 1 ~ "NE", 
                                response_cat == 2 ~ "both", 
                                T ~ "E")) %>% 
      mutate(response = factor(response, levels = c("NE", "both", "E")))
  
  df.obs <- dat %>% filter(stimulus == !! stimulus_letter) %>% 
    group_by(QUD, response_int, response) %>% summarize(n=n(), .groups = "drop_last") %>% 
    group_by(QUD) %>% mutate(p = n / sum(n)) 
    
  p <- pp_samples %>%
    ggplot(aes(x=p, color = QUD)) + geom_density() +
    facet_wrap(~response, scales = "free_y") + 
    geom_point(data=df.obs, aes(x=p, y=0)) 
  return(p)
}
plot_posterior_check_by_stimulus(df.ordinal, "A") + ggtitle("stimulus: A")
plot_posterior_check_by_stimulus(df.ordinal, "C") + ggtitle("stimulus: C")

plot_posterior_check_by_stimulus(df.ordinal, "B") + ggtitle("stimulus: B")
plot_posterior_check_by_stimulus(df.ordinal, "D") + ggtitle("stimulus: D")
```



### Exploratory analysis

```{r}
# H2: P(E | QUD=willq) > P(both | QUD=willq)
samples_posterior_probs %>% filter(QUD == "willq") %>% mutate(h2 = E > both) %>%
  group_by(stimulus) %>% summarize(p_h2 = mean(h2)) %>% 
  mutate(p_h2.across = mean(p_h2))

# H3: P(both | QUD=ifp) > P(E | QUD=ifp)
samples_posterior_probs %>% filter(QUD == "ifp") %>% mutate(h3 = both > E) %>%
  group_by(stimulus) %>% summarize(p_h3 = mean(h3)) %>% 
  mutate(p_h3.across = mean(p_h3))
```

For stimulus A and C P(both | QUD=ifp) > P(E | QUD=ifp) is 1. 
 
 - What about P(E | QUD=willq) > P(both | QUD=willq)? (stimuli B+D)
 - further, is  P(E | QUD=ifp) > P(NE | QUD=ifp)

```{r}
samples_posterior_probs %>% filter(QUD == "willq") %>% 
  mutate(h3_willq = E > both) %>% group_by(stimulus) %>% 
  summarize(p_h3_willq = mean(h3_willq)) %>% 
  mutate(p_h3_willq.across = mean(p_h3_willq))

samples_posterior_probs %>% filter(QUD == "ifp") %>% 
  mutate(h = E > NE) %>% group_by(stimulus) %>% 
  summarize(p_h = mean(h)) %>% 
  mutate(p_h.across = mean(p_h))
```

Check P(NE) for QUD:if-p vs. QUD:will-q

```{r}
samples_posterior_probs %>% 
  filter(stimulus == "B" | stimulus == "D") %>% 
  dplyr::select(.draw, QUD, stimulus, NE) %>% 
  group_by(QUD, stimulus) %>% 
  pivot_wider(names_from = "QUD", values_from = "NE") %>% 
  mutate(h_NE = willq > ifp) %>% 
  summarize(h_NE = mean(h_NE))

# P(NE) close to 0? Not for stimulus C!
samples_posterior_probs %>% 
  dplyr::select(.draw, stimulus, NE) %>% 
  group_by(stimulus) %>% 
  mutate(h_NE = NE <= 0.1) %>% 
  summarize(h_NE = mean(h_NE))
```

Difference between stimulus B and D?

```{r}
df = samples_posterior_probs %>% filter(stimulus == "B" | stimulus == "D") %>% 
  dplyr::select(.draw, QUD, stimulus, E, both, NE) %>% 
  pivot_longer(cols=c(NE, both, E), names_to = "response", values_to = "p") %>% 
  group_by(QUD, stimulus) 

# both if-p
df %>% filter(QUD == "ifp" & response == "both") %>% 
  pivot_wider(names_from = "stimulus", values_from = "p") %>% 
  mutate(h = B < D) %>% summarize(p_h = mean(h))
# exhaustive if-p
df %>% filter(QUD == "ifp" & response == "E") %>% 
  pivot_wider(names_from = "stimulus", values_from = "p") %>% 
  mutate(h = B > D) %>% summarize(p_h = mean(h))
# same with hypothesis-function: is B more exhaustive when QUDif-p than D?
hypothesis(ordinal_model, "exhwoD > 0")


# both will-q
df %>% filter(QUD == "willq" & response == "both") %>% 
  pivot_wider(names_from = "stimulus", values_from = "p") %>% 
  mutate(h = B < D) %>% summarize(p_h = mean(h))
# Exhaustive will-q
df %>% filter(QUD == "willq" & response == "E") %>% 
  pivot_wider(names_from = "stimulus", values_from = "p") %>% 
  mutate(h = B > D) %>% summarize(p_h = mean(h))
# same with hypothesis function: is B more exhaustive when QUD:will-q than D?
hypothesis(ordinal_model, "QUDwillMq + exhwoD + QUDwillMq:exhwoD > QUDwillMq")

```
Difference between stimuli A and C?

```{r}
df = samples_posterior_probs %>% filter(stimulus == "A" | stimulus == "C") %>% 
  dplyr::select(.draw, QUD, stimulus, E, both, NE) %>% 
  pivot_longer(cols=c(NE, both, E), names_to = "response", values_to = "p") %>% 
  group_by(QUD, stimulus) 

# NE more often selected for stimulus C than for A?
df %>% filter(QUD == "ifp" & response == "NE") %>% 
  pivot_wider(names_from = "stimulus", values_from = "p") %>% 
  mutate(h = C > A) %>% summarize(p_h = mean(h))
```


Model summary: only main effect b_nonExhint does not include 0 in 95% interval

```{r}
summary(ordinal_model)
mcmc_areas(ordinal_model, regex_pars = "b_")
```


##  Reaction times

Look at reaction times split across all picture pairs

```{r}
N=100
n_participants = data %>% distinct_at(vars(c(submission_id))) %>% nrow()
critical.RT = data.critical %>% 
  dplyr::select(submission_id, QUD, RT, pic_condition, selected_pic, stimulus) %>%
  mutate(visual_diff = case_when(stimulus %in% c("B", "C") ~ T, 
                                 T ~ F)) %>% 
  group_by(QUD, pic_condition) 
critical.RT_means = critical.RT %>% 
  summarize(mean_RT = mean(RT), .groups = "drop_last")

bootstrap_RTs = group_map(critical.RT, function(df, df.group){
  bootstrap = map(seq(1, N), function(i){
    tibble(mean_RT = mean(sample(df$RT, n_participants, replace=T)))
  }) %>% bind_rows() %>% 
    add_column(pic_condition = df.group$pic_condition, QUD = df.group$QUD) %>% 
    arrange(mean_RT)
  
  N = bootstrap %>% nrow()
  CI.bounds = c(low=ceiling(0.025 * N), up=ceiling(0.975 * N))

  ci.low = bootstrap[CI.bounds[["low"]], ]$mean_RT
  ci.up = bootstrap[CI.bounds[["up"]], ]$mean_RT
  bootstrap %>% add_column(ci.low=ci.low, ci.up=ci.up)
}) %>% bind_rows() %>% group_by(pic_condition, QUD)

bootstrap_RTs.CI.question_pic = bootstrap_RTs %>%
  dplyr::select(-mean_RT) %>% distinct()

df <-left_join(critical.RT_means, bootstrap_RTs.CI.question_pic) %>% 
  separate(pic_condition, into=c("exhaustive", "non_exhaustive"), sep="-") %>% 
  mutate(exhaustive = case_when(exhaustive == "w/oDistractor" ~ "woD", T ~ "withD"), 
         non_exhaustive = case_when(non_exhaustive == "internal" ~ "int", T ~ "ext")) %>%
  map_conditions_to_stimuli()
  
p <- df %>% 
  ggplot(aes(y=mean_RT, x=exhaustive, color=non_exhaustive, shape = QUD)) + 
  geom_point(stat="identity", position = position_dodge(width=0.25)) + 
  geom_line(aes(group=interaction(QUD, non_exhaustive)),
            position = position_dodge(width=0.25)) +
  geom_text(aes(label = stimulus), vjust=0.9) + 
  geom_errorbar(aes(ymin=ci.low, ymax=ci.up), position = position_dodge(width=0.25))+
  facet_wrap(~QUD)
p
ggsave(file.path(plot_dir, "reaction-times-233.pdf"), p, width=7, height=4)

```

```{r}
p.rt = critical.RT %>%
  ggplot(aes(x=log(RT), color=pic_condition)) + geom_density() +
  guides(color = guide_legend(ncol=2))  +
  scale_color_brewer(name = "stimulus", palette = "Set2")

p.rt
p.rt + facet_wrap(~QUD)
ggsave(file.path(plot_dir, "empiric-rts.pdf"), p.rt, width=7, height=4)
```



