---
title: "analysis"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(boot)
library(brms)
library(here)
library(latex2exp)
library(scales)
library(tidybayes)
library(tidyverse)
source("utils.R")

theme_set(theme_minimal(base_size=12) + theme(legend.position = "top"))
target_dir = here("results", "pretests")
# target_dir = here("results")

plot_dir = file.path(target_dir, "figs")
if(!file.exists(plot_dir)) dir.create(plot_dir, recursive = TRUE)
```

# Load cleaned experimental data

Data was cleaned (see script R/clean_data.R) according to predefined exclusion criteria

```{r, message = FALSE}
data = read_csv(here("results", "data_cleaned.csv")) %>% group_by(prolific_id)
data.test = data %>% filter(str_detect(trial_name, "image_selection"))
```

## Participants' responses to questions shown in the end

```{r}
data %>% dplyr::select(submission_id, check_ann) %>% 
  distinct() %>% group_by(check_ann) %>% dplyr::count()
data %>% dplyr::select(submission_id, check_behavior) %>% 
  distinct() %>% group_by(check_behavior) %>% dplyr::count()
```

## Trials where a pragmatic and a literal selection is possible

```{r}
trials_lit_prag = c("trial7", "trial8", "trial19", "trial20", "trial21", "trial24")
data.lit_prag <- data %>% filter(id %in% trials_lit_prag) %>% 
  dplyr::select(submission_id, response, type, id, response) %>% group_by(id)

# by trials
data.lit_prag %>% 
  ggplot(aes(x=id, fill=response)) + 
  geom_bar(stat="count", position=position_dodge()) +
  scale_y_continuous(breaks = pretty_breaks())

pic_props = data.test %>% ungroup() %>% 
  select(starts_with("property_"), type, id) %>% 
  unite(col="props", "property_pic1", "property_pic2", "property_pic3", sep="_") %>% 
  distinct() %>% arrange(type)

pic_props %>% filter(id %in% trials_lit_prag)
```

## "Pragmatic" scores
Scores for participants how pragmatic their responses were. 
1: pragmatic picture selected (pragmatic)
0.5: pragmatic + literal picture selected (literal2)
0: literal picture selected (literal1)

Max score:6
Min score:0

```{r}
data.pragmatic_score = data.lit_prag %>% 
  mutate(score = case_when(response == "pragmatic" ~ 1,
                           response == "literal" ~ 0, 
                           T ~ 0.5)) %>% 
  group_by(submission_id) %>% summarize(score = sum(score), .groups = "drop") %>% 
  mutate(relative_score = round(score / 6, 2))

data.pragmatic_score %>% 
  ggplot(aes(x=score)) + geom_bar() + 
  labs(x = "pragmatic score", y = "nb participants") +
  scale_x_continuous(breaks = pretty_breaks())
```

## Practice trials

```{r}
data.practice = data.test %>% filter(type == "practice") %>% 
   dplyr::select(submission_id, response, id)

data.practice %>% ggplot(aes(x=id, fill=response)) + 
  geom_bar(stat="count", position=position_dodge()) +
  scale_y_continuous(breaks = pretty_breaks())

pic_props %>% filter(type == "practice")
```

## Trials where we'd expect selection of exhaustive + non-exhaustive

Trials, where a selection of one is risky since no clue for neither of both
Trial11 is a practice trial, trial13 a filler trial.

```{r}
risk_trials = c("trial11", "trial13")
data.risk <- data %>% filter(id %in% risk_trials) %>% 
  dplyr::select(submission_id, response, id, type)

data.risk %>% ggplot(aes(x=id, fill=response)) + 
  geom_bar(stat="count", position=position_dodge()) +
  scale_y_continuous(breaks= pretty_breaks())

pic_props %>% filter(id %in% risk_trials)
```

## Meta information about participants

```{r, meta, echo=FALSE}
data.meta <- data %>% 
  distinct_at(vars(c(submission_id)), .keep_all = T) %>%
  mutate(education=as_factor(education), gender=as_factor(gender))

data.meta$gender %>% summary()
data.meta$education %>% summary()
print("age:")
data.meta$age %>% summary()
print("timeSpent:")
data.meta$timeSpent %>% summary()

data.meta %>% ggplot(aes(x="time", y=timeSpent)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter() +
  labs(x="")
```

## Participants' game scores from test phase

Horizontal lines are maximal scores that can be reached in respective trials

```{r}
data.test_scores = data %>% filter(type != "training") %>% 
  dplyr::select(prolific_id, submission_id, score, type, id) %>%
  group_by(type, prolific_id) %>%
  mutate(prolific_id = as.factor(prolific_id))

df.max = data.test_scores %>% group_by(type) %>% dplyr::select(type, id) %>% 
  distinct() %>% summarize(n=n()) %>% 
  mutate(max = case_when(type == "practice" ~ n*100-50, 
                         T ~ n * 100))
data.test_scores %>% 
  summarize(score = sum(score), .groups = "drop_last") %>% 
  ggplot(aes(x=prolific_id, y=score, fill=type)) + 
  geom_hline(data = df.max, aes(yintercept = max, color = type)) +
  geom_bar(stat="identity", position = position_dodge()) +
  facet_wrap(~prolific_id, scales = "free")
```

## Training data

Sanity check: does ratio of correct answers increase with increasing trial numbers?

```{r}
data.train = data %>% filter(type=="training") %>%
  mutate(correct=case_when(expected == response ~ T, T ~ F))

# ratio correct by participants
data.train %>% dplyr::select(prolific_id, id, correct) %>% 
  summarize(nb_trials = n(), nb_correct = sum(correct)) %>% 
  mutate(ratio_correct = nb_correct / nb_trials) %>% 
  ggplot(aes(x=prolific_id, y=ratio_correct)) + geom_bar(stat="identity") +
  geom_hline(aes(yintercept = 0.5))

# ratio correct per trial number (different order per participant)
train.correct.nb = data.train %>% group_by(trial_number) %>% 
  summarize(mean_correct = mean(correct))
train.correct.nb %>% 
  ggplot(aes(x=trial_number, y=mean_correct)) + 
  geom_bar(stat="identity") + 
  scale_x_continuous(breaks = scales::pretty_breaks(8)) +
  geom_hline(aes(yintercept = 0.5))

# ratio correct per trial id
train.correct.id = data.train %>% group_by(id) %>% 
  summarize(mean_correct = mean(correct))
train.correct.id %>% 
  ggplot(aes(x=id, y=mean_correct)) + 
  geom_bar(stat="identity") +
  theme(axis.text.x=element_text(angle=90)) + 
  geom_hline(aes(yintercept = 0.5))
```

# Critical test trials

Plot critical trials with 95%-bootstrap confidence intervals

```{r, data-critical, echo=FALSE}
data.critical = data %>% 
  filter(type=="critical") %>% 
  dplyr::select(submission_id, prolific_id, id, type, QUD, response, 
                selected_pic, pic2, pic3, RT) %>% 
  mutate(pic2=str_replace(pic2, ".png", ""), pic2=str_replace(pic2, "_horiz", ""),
         pic3 = str_replace(pic3, ".png", ""), pic3=str_replace(pic3, "_horiz", "")) %>% 
  mutate(picPair=paste(pic2, pic3, sep=" & "),
         pic_condition = 
           case_when(picPair == "if2_unn & if2_unu" ~ "D. withDistractor-external", 
                     picPair == "if1_un & if2_unu" ~ "B. w/oDistractor-external",
                     picPair == "if1_un & if1_uu" ~ "A. w/oDistractor-internal",
                     picPair == "if2_unn & if1_uu" ~ "C. withDistractor-internal")) %>% 
  dplyr::select(-pic2, -pic3) %>%
  mutate(stimulus = substr(pic_condition, 1, 1),
         pic_condition = substr(pic_condition, 4, nchar(pic_condition)), 
         response = case_when(response == "exhaustive_non-exhaustive" ~ "both",
                              T ~ response),
         QUD = case_when(QUD == "willq" ~ "will-q",
                         QUD == "ifp" ~ "if-p", 
                         T ~ QUD))
```

```{r}
df.critical <- data.critical %>% 
  group_by(pic_condition, stimulus, QUD) %>% 
  dplyr::select(submission_id, pic_condition, stimulus, QUD, response)

df.crit.exh = df.critical %>% 
  mutate(response = case_when(response == "exhaustive" ~ 1, T ~ 0))
df.crit.nonExh = df.critical %>% 
  mutate(response = case_when(response == "non-exhaustive" ~ 1, T ~ 0))
df.crit.both = df.critical %>% 
  mutate(response = case_when(response == "both" ~ 1, T ~ 0))
df.crit.other = df.critical %>% 
  mutate(response = case_when(!response %in% c("exhaustive", "non-exhaustive", 
                                               "both") ~ 1, T ~ 0))
bootstrap_samples = bind_rows(
  get_bootstrap_statistics(df.crit.both, "both"),
  get_bootstrap_statistics(df.crit.exh, "exhaustive"),
  get_bootstrap_statistics(df.crit.nonExh, "non-exhaustive"),
  get_bootstrap_statistics(df.crit.other, "other selection")
) %>% group_by(pic_condition, stimulus, QUD, response)
CIs <- get_bootstrap_cis(bootstrap_samples)

data.critical.rate = df.critical %>% 
  group_by(pic_condition, stimulus, QUD, response) %>% 
  summarize(n=n(), .groups = "drop_last") %>% 
  mutate(ratio = n/sum(n))

data.critical.rate = left_join(data.critical.rate, CIs, 
  by=c("QUD", "pic_condition", "stimulus", "response")) %>% 
  mutate(pic_condition = factor(pic_condition, 
                              levels = c("w/oDistractor-internal", 
                                         "w/oDistractor-external", 
                                         "withDistractor-internal",
                                         "withDistractor-external")), 
         response = factor(response, levels = c("non-exhaustive", "both",
                                              "exhaustive")),
         QUD = as.factor(QUD))
```

## Plotted across all different picture combinations

```{r}
QUDs = list(neutral = "Which blocks do you think will fall?",
            `if-p` = "What happens if the 'antecedent-block' falls?",
            `will-q` = "Will the 'consequent-block' fall?")
data.critical %>%
  mutate(response = factor(response, levels = c("non-exhaustive", "both",
                                                "exhaustive"))) %>% 
  ggplot(aes(fill=response, y=QUD)) + 
  geom_bar(position=position_dodge(preserve = 'single')) +
  scale_y_discrete(breaks=names(QUDs), labels = QUDs) + 
  labs(y="Ann's question") +
  guides(fill = guide_legend(ncol = 2))

# same plot with exhaustive-picture selection ratios instead of absolute numbers
data.critical.rate %>% group_by(QUD, response) %>% 
  mutate(ratio = mean(ratio)) %>% 
  distinct_at(vars(c(QUD, response)), .keep_all = T) %>% 
  ggplot(aes(x=ratio, y=QUD, fill=response, group = response)) + 
  geom_bar(stat="identity", position = position_dodge()) + 
  geom_errorbar(aes(xmin=ci.low, xmax=ci.up), 
                position = position_dodge()) +
  geom_text(aes(label=round(ratio, 2)), hjust=-0.25, vjust=-0.45, color = 'blue',
            position = position_dodge(width = .9)) +
  scale_y_discrete(breaks=names(QUDs), labels = QUDs) + 
  labs(y="Ann's question", x="selection rate") +
  guides(fill = guide_legend(ncol = 2))

# point plot
w_dodge = 0.25
p = data.critical.rate %>% group_by(QUD, response) %>% 
  mutate(ratio = mean(ratio)) %>% 
  distinct_at(vars(c(QUD, response)), .keep_all = T) %>% 
  ggplot(aes(y=ratio, x=QUD, color=response, group=response)) + 
  geom_point(position = position_dodge(width=w_dodge)) +
  geom_line(position = position_dodge(width = w_dodge)) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.up), width=0.1, 
                position = position_dodge(width=w_dodge)) +
  labs(y="selection rate", x = "QUD")
p
```

## plotted separately for each picture combination

```{r}
p = data.critical.rate %>% 
  ggplot(aes(x=ratio, y=QUD, fill=response, group=response)) + 
  geom_bar(stat="identity", position = position_dodge()) + 
  geom_errorbar(aes(xmin=ci.low, xmax=ci.up), position = position_dodge()) +
  geom_text(aes(label=round(ratio, 2)), hjust=-0.25, vjust=-0.3,
            position = position_dodge(width = .9), color='blue') +
  facet_wrap(~pic_condition) + 
  labs(x="selection rate", y = "QUD")
p
ggsave(here(plot_dir, "selection-rates-barplot.pdf"), p, width=7, height=4)

w_dodge = 0.25
p = data.critical.rate %>% 
  ggplot(aes(y=ratio, x=QUD, color=response, group=response)) + 
  geom_point(position = position_dodge(width=w_dodge)) +
  geom_line(position = position_dodge(width = w_dodge)) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.up), width=0.1, 
                position = position_dodge(width=w_dodge)) +
  facet_wrap(~pic_condition) + 
  labs(y="selection rate", x = "QUD")
p
ggsave(here(plot_dir, "selection-rates.pdf"), p, width=7, height=4)
```

##  Responses for QUD:will-q and QUD:if-p plotted together

```{r}
get_responses_all_QUDs = function(dat){
  df = dat %>% group_by(submission_id, pic_condition, stimulus) %>% 
    dplyr::select(submission_id, pic_condition, stimulus, QUD, response)

  df.wide = df %>% group_by(pic_condition, stimulus, submission_id) %>% 
    mutate(response = str_replace(response, "non-exhaustive", "NE"), 
           response = str_replace(response, "exhaustive", "E")) %>% 
    pivot_wider(names_from = QUD, values_from = response, values_fill = "-") %>% 
    unite("ifp.willq", `if-p`, `will-q`, sep="/") %>% 
    separate(pic_condition, into = c("exh", "nonExh"), sep = "-") %>%
    rename(response = ifp.willq)
  return(df.wide)
}

plot_responses_all_QUDs = function(dat) {
  df.wide <- get_responses_all_QUDs(dat)
  df.wide.count = df.wide %>% group_by(stimulus, exh, nonExh, response) %>% 
    dplyr::count() %>% group_by(stimulus) %>% mutate(N=sum(n), ratio=n/N)
  
  xlab = "selected situations for both QUDs: if-p/will-q"
  p <- df.wide.count %>% ggplot(aes(x=response, y=ratio , fill=response)) + 
    geom_bar(stat="identity", position=position_dodge()) +
    geom_label(aes(label=n)) +
    scale_fill_brewer(palette='Set2', name="response ifp_willq") +
    facet_wrap(~stimulus, nrow = 2, scales = "free") +
    theme(legend.position = "none") +
    labs(y = "ratio participants", x=xlab)
  return(p)
}
p <- plot_responses_all_QUDs(data.critical)
p
ggsave(file.path(plot_dir, "empirical-responses.pdf"), p, width=7, height=4)
```

## Scores concerning the nb. of times participants give different responses depending on QUD

Score participants: for the two critical stimuli that occur with both QUDs (B +
D), participants receive 1 point if they differentiate the QUDs ifp_willq, i.e.
when they choose the exhaustive picture for one QUD and the non-exhaustive for
the other (independent of direction). A score of 0 then means that the
respective participant never made a difference between the two QUDs and a score
of two that participant always made a difference between them.

```{r}
responses.all_quds = get_responses_all_QUDs(data.critical) %>% 
  separate("response", into = c("ifp", "willq"), sep="/") %>% 
  mutate(score = case_when(ifp == willq ~ 0, T ~ 1)) %>%
  filter(willq != "-") # some stimuli are not included with QUD=will-q

df.qud_scores = responses.all_quds %>% group_by(submission_id) %>%
  summarize(score.total=sum(score))
df.qud_scores.summary = df.qud_scores %>% group_by(score.total) %>% 
  dplyr::count() %>% ungroup() %>% mutate(N=sum(n)) %>% 
  mutate(ratio = n/N)

df.qud_scores.summary %>% ggplot(aes(x=score.total, y=ratio)) + 
  geom_bar(stat="identity") +
  labs(x="nb times participants gave different answers for different QUDs", 
       y = "nb participants")

ids.scores12 <- df.qud_scores %>% filter(score.total <= 2 & score.total >= 1) %>% 
  pull(submission_id)
plot_responses_all_QUDs(data.critical %>% filter(submission_id %in% ids.scores12)) +
  ggtitle("Participants with 0 or 1 ignorance of QUD")
```

Those who never take QUD into account, do they switch between exh-exh and 
nonExh-nonExh or always give the same (identical) answer?

```{r}
ids.score0 = df.qud_scores %>% filter(score.total == 0) %>% pull(submission_id)
df.ignored_qud = responses.all_quds %>% filter(submission_id %in% ids.score0) %>% 
  group_by(submission_id, ifp, willq) %>% 
  dplyr::count(name="nb_stimuli") %>%
  arrange(submission_id) %>% 
  rename(selection = ifp) %>% ungroup() %>% dplyr::select(-willq) %>% 
  mutate(switch_answer = case_when(nb_stimuli == 2 ~ FALSE, TRUE ~ TRUE))
df.ignored_qud

# plot QUD-score from test trials against pragmatic score from practice trials
# pair with participants pragmatic scores
scores = left_join(df.qud_scores %>% rename(score_qud = score.total) %>% 
                     mutate(score_qud = as.integer(score_qud)), 
                   data.pragmatic_score %>% 
                     dplyr::select(submission_id, relative_score) %>% 
                     rename(score_pragmatic = relative_score), 
          by="submission_id")

scores %>% 
  mutate(submission_id = as.factor(submission_id), 
         score_qud = as.factor(score_qud)) %>% 
  ggplot(aes(x=score_qud, y=score_pragmatic)) +
  geom_jitter(width=0.2) + theme(legend.position = "none")

```

## Plot empirical selection rates

```{r}
df <- data.critical.rate %>% 
  separate(col = pic_condition, into = c("exh_situation", "nonExh_situation"), sep="-") %>%
  mutate(exh_situation = factor(exh_situation, levels = c("withDistractor", "w/oDistractor")),
         response = str_replace(response, "non-exhaustive", "nonExh"),
                    response = str_replace(response, "exhaustive", "exh"))

w_dodge = 0.25

p1 <- df %>%
  ggplot(aes(x=exh_situation, y=ratio, color = response, shape = stimulus)) +  
             # group=interaction(QUD, nonExh_situation))) +
  geom_point(position = position_dodge(width=w_dodge)) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.up), width=0.1, 
                position = position_dodge(width=w_dodge)) +
  # geom_line(position = position_dodge(width=w_dodge)) +
  # labs(y = "selection rates", x = "exhaustive situation") +
  scale_color_brewer(name = "selection", palette = "Set2") +
  guides(color = guide_legend(title.position="top"), 
         shape = guide_legend(title = "stimulus", title.position="top")) +
  facet_grid(QUD ~ nonExh_situation, scales = "free_x", labeller = label_both)
p1
ggsave(file.path(plot_dir, "empirical-responses-rates-fig1.pdf"), p1, 
       width=7, height=4)

p2 <- df %>% 
  ggplot(aes(x=stimulus, y=ratio, color = response)) +
             #group=interaction(QUD, nonExh_situation))) +
  geom_point(position = position_dodge(width=w_dodge)) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.up), width=0.1, 
                position = position_dodge(width=w_dodge)) +
  #geom_line(position = position_dodge(width=w_dodge)) +
  #labs(y = "selection rates", x = "exhaustive situation") +
  labs(x="stimulus", y="selection rate") +
  scale_color_brewer(name = "selection", palette = "Set2") +
  #guides(color = guide_legend(title.position="top")) +
         #shape = guide_legend(title = "stimulus", title.position="top")) +
  facet_grid(QUD ~ stimulus, scales = "free", labeller = label_both)

p2
ggsave(file.path(plot_dir, "empirical-responses-rates-fig2.pdf"), p2, 
       width=7, height=4)

p3 <- df %>% 
  ggplot(aes(x=QUD, y=ratio, color = response, group = response)) +
             #group=interaction(QUD, nonExh_situation))) +
  geom_point(position = position_dodge(width=w_dodge)) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.up), width=0.1, 
                position = position_dodge(width=w_dodge)) +
  geom_line(position = position_dodge(width=w_dodge)) +
  labs(x="stimulus", y="selection rate") +
  scale_color_brewer(name = "selection", palette = "Set2") +
  facet_grid(~ stimulus, scales = "free", labeller = label_both)

p3
ggsave(file.path(plot_dir, "empirical-responses-rates-fig3.pdf"), p2, 
       width=7, height=4)

```


# Bayesian regression models

Prepare Data

```{r}
data.critical.model = data.critical %>% 
  dplyr::select(-picPair, -selected_pic, -submission_id, -type, -id) %>% 
  mutate(picPair = pic_condition, 
         nb_selected = str_count(response, "_") + 1,
         response = case_when(nb_selected == 3 ~ "all", T ~ response),
         response = str_replace(response, "non-exhaustive", "NE"),
         response = str_replace(response, "exhaustive", "E"), 
         response = str_replace(response, "contrast", "C"), 
         response = str_replace(response, "E_NE", "both"), 
         QUD = as.factor(QUD)) %>% 
  separate(col = pic_condition, into = c("exh", "nonExh"), sep="-") %>% 
  mutate(exh = case_when(exh == "w/oDistractor" ~ "woD", 
                         exh == "withDistractor" ~ "withD"),
         exh = factor(exh, levels = c("woD", "withD")), 
         nonExh = case_when(nonExh == "internal" ~ "int", 
                            nonExh == "external" ~ "ext"),
         nonExh = factor(nonExh, levels = c("int", "ext")))
```


## Bayesian ordinal regression model

###  Main analysis 

```{r}
df.ordinal <- data.critical.model %>%
  mutate(response = factor(response, levels = c("NE", "both", "E"), ordered=T), 
         exh = factor(exh, levels = c("withD", "woD")),
         nonExh = factor(nonExh, levels = c("ext", "int")), 
         QUD = factor(QUD, levels = c("if-p", "will-q"))
         )
contrasts(df.ordinal$response) <- contr.treatment(3)

ordinal_model <- brm(data = df.ordinal,
  family = cumulative("probit"),
  formula =  response ~ 1 + QUD * exh + exh * nonExh + 
    (1 + exh + nonExh + QUD | prolific_id),
  seed = 1, chains = 4, cores = 4, iter = 3000)

conds <- make_conditions(df.ordinal, c("exh", "nonExh"))
effects <- conditional_effects(ordinal_model, "QUD", categorical = TRUE, 
                               conditions = conds, robust = T, 
                               re_formula = NA)
# effects$`QUD:cats__`
effects

summary(ordinal_model)

hypothesis(ordinal_model, "QUDwillMq > 0")

```

### Exploratory analysis

```{r}
# Thresholds
tau1 = fixef(ordinal_model)["Intercept[1]", "Estimate"]
tau2 = fixef(ordinal_model)["Intercept[2]", "Estimate"]
posterior_draws = tidy_draws(ordinal_model)

# reference category: QUD:if-p, exh:withD, nonExh:ext
posterior_preds = function(qud, exh, non_exh, tau1=tau1, tau2=tau2, 
                           draws=posterior_draws) {

  qud = str_replace(qud, "-", "") # to catch if-p/will-q inputs as well
  samples <- draws %>% dplyr::select(starts_with("b_")) %>% 
    rename(tau1 = `b_Intercept[1]`, tau2 = `b_Intercept[2]`) %>% 
    mutate(
      eta_willq_withD_ext = `b_QUDwillMq`,
      eta_ifp_withD_int = `b_nonExhint`,
      eta_ifp_woD_ext = `b_exhwoD`,
      
      eta_ifp_woD_int = `b_exhwoD` + `b_nonExhint` + `b_exhwoD:nonExhint`,
      eta_willq_woD_ext = `b_QUDwillMq` + `b_exhwoD` + `b_QUDwillMq:exhwoD`
    )
  eta = paste("eta", qud, exh, non_exh, sep = "_")
  
  if(eta == "eta_ifp_withD_ext") {
    # predictions for reference category
    predictions = samples %>% 
      transmute(NE = pnorm(tau1), both = pnorm(tau2) - pnorm(tau1), 
                E = 1 - pnorm(tau2))
  } else {
    predictions = samples %>% 
      transmute(NE = pnorm(tau1 - .data[[eta]]), 
                both = pnorm(tau2 - .data[[eta]]) - 
                  pnorm(tau1 - .data[[eta]]),
                E = 1 - pnorm(tau2 - .data[[eta]])) 
  }
  
  predictions %>% add_column(QUD=qud, exhaustive=exh, non_exhaustive=non_exh)
}

samples_posterior_probs =
  bind_rows(posterior_preds("willq", "withD", "ext"),
            posterior_preds("willq", "woD", "ext"),
            posterior_preds("ifp", "withD", "ext"),
            posterior_preds("ifp", "woD", "ext"),
            posterior_preds("ifp", "withD", "int"),
            posterior_preds("ifp", "woD", "int"))


# H2: P(E | QUD=willq) > P(both | QUD=willq)
samples_posterior_probs %>% filter(QUD == "willq") %>% 
  mutate(h2 = E > both) %>% summarize(p_h2 = mean(h2))

# H3: P(both | QUD=ifp) > P(E | QUD=ifp)
samples_posterior_probs %>% filter(QUD == "ifp") %>% 
  mutate(h3 = both > E) %>% summarize(p_h3 = mean(h3))
```

## Bayesian logistic regression model

```{r, results='hide'}
model_path <- here(target_dir, "brms-model-QUD-exh-nonExh.rds")

if(file.exists(model_path)) {
  model <- readRDS(model_path)
} else {
model <-
  brm(data = data.critical.model,
      family = categorical(link="logit"),
      response ~ 1 + QUD * exh * nonExh + (1 + QUD * exh * nonExh| prolific_id),
      seed = 1, chains = 4, cores = 4, iter = 4000)
  saveRDS(model, file=model_path)
}
model_params <- fixef(model)[, 'Estimate']
pp_check(model)
```


```{r}
# function that transforms linear predictors on odds-scale into probabilities
transform_to_probs = function(predictors_grouped, ref_cat = "both", simplify = F){
  df.out <- predictors_grouped %>% 
    mutate(denominator = sum(exp(value)) + 1,
           p = exp(value)/denominator,
           #p_ref_cat = 1-sum(p), # is the same as
           p_ref = 1/denominator, # this
           ref_cat = ref_cat) %>% 
    filter(!is.infinite(denominator)) # QUD:willq not used for 2 stimuli
  
  if(simplify){
    df = df.out %>% dplyr::select(-value, -denominator)
    df.ref_cat <- df %>% dplyr::select(-p, -prediction) %>% 
      rename(p=p_ref, prediction = ref_cat) %>% distinct()
    df.out = bind_rows(df %>% dplyr::select(-p_ref, -ref_cat),
                       df.ref_cat)
  }
  return(df.out)
}
```

- reference output category (Intercept): both
- reference level for predictor QUD: if-p
- reference level for predictor exhaustive: woD
- reference level for predictor non-exhaustive: int (internal)

Some predicted probabilities

```{r}
get_linear_predictors = function(predicted_cat, params) {
  mu = paste('mu', predicted_cat, sep='')
  pars = params[startsWith(names(params), mu)] 
  df = tibble(
    prediction = predicted_cat,
    # stimulus A:
    ifp_woD_int = pars[[paste(mu, '_Intercept', sep='')]],
    willq_woD_int = ifp_woD_int + pars[[paste(mu, '_QUDwillq', sep='')]],
    # stimulus B:
    ifp_woD_ext = ifp_woD_int + pars[[paste(mu, '_nonExhext', sep='')]],
    willq_woD_ext = ifp_woD_int + pars[[paste(mu, '_nonExhext', sep='')]] + 
      pars[[paste(mu, '_QUDwillq', sep='')]] + 
      pars[[paste(mu, '_QUDwillq:nonExhext', sep='')]],
    # stimulus C:
    ifp_withD_int = ifp_woD_int + pars[[paste(mu, '_exhwithD', sep='')]],
    willq_withD_int = ifp_woD_int + pars[[paste(mu, '_exhwithD', sep='')]] +
      pars[[paste(mu, '_QUDwillq', sep='')]] +
      pars[[paste(mu, '_QUDwillq:exhwithD', sep='')]],
    # stimulus D: 
    ifp_withD_ext = ifp_woD_int + pars[[paste(mu, '_exhwithD', sep='')]] + 
      pars[[paste(mu, '_nonExhext', sep='')]] + 
      pars[[paste(mu, '_exhwithD:nonExhext', sep='')]],
    willq_withD_ext = ifp_woD_int + pars[[paste(mu, '_exhwithD', sep='')]] + 
      pars[[paste(mu, '_nonExhext', sep='')]] + 
      pars[[paste(mu, '_QUDwillq', sep='')]] + 
      pars[[paste(mu, '_exhwithD:nonExhext', sep='')]]+
      pars[[paste(mu, '_QUDwillq:exhwithD', sep='')]] +
      pars[[paste(mu, '_QUDwillq:nonExhext', sep='')]] +
      pars[[paste(mu, '_QUDwillq:exhwithD:nonExhext', sep='')]]
  )
  return(df)
}

# linear predictors
predictors_E = get_linear_predictors("E", model_params)
predictors_NE = get_linear_predictors("NE", model_params)

# transformed to probabilities
predictors = bind_rows(predictors_E, predictors_NE)
df <- predictors %>% pivot_longer(cols=c(-prediction), names_to = "param") %>% 
    separate(param, into=c("QUD", "exh", "non_exh")) %>% group_by(QUD, exh, non_exh)
predicted_p = transform_to_probs(df, simplify = T)
predicted_p %>% arrange(QUD, exh, non_exh, prediction)
```

Plot conditional effects on probability scale: 

```{r}
conds <- make_conditions(data.critical.model, c("exh", "nonExh"))
p = conditional_effects(
  model, effects = c("QUD"), conditions = conds, categorical = T, 
  re_formula = NA, method = "fitted", robust = F
)
plot(p, facet_args = c(nrow = 2, ncol=2, labeller =
                         as_labeller(c(`exh = woD & nonExh = int` = "stimulus A", 
                                       `exh = woD & nonExh = ext` = "stimulus B",
                                       `exh = withD & nonExh = int` = "stimulus C",
                                       `exh = withD & nonExh = ext` = "stimulus D"))))
cols = c("cats__", "QUD", "exh", "nonExh", "estimate__", "se__", "lower__", "upper__")
predicted_vals = p[["QUD:cats__"]][, cols] %>% tibble()
predicted_vals
```

Test hypotheses

```{r}
posterior_draws = tidy_draws(model)
posterior_draws_probs = function(df) {
  samples = df %>% as_tibble() %>% dplyr::select(starts_with("b_mu")) %>% 
    pivot_longer(cols = everything()) %>%
    mutate(name = str_replace(name, "b_", "")) %>% deframe()
 
  predictors = bind_rows(get_linear_predictors("E", samples),
                         get_linear_predictors("NE", samples)) %>% 
    pivot_longer(cols=c(-prediction), names_to = "param") %>% 
    separate(param, into=c("QUD", "exh", "nonExh"), sep="_") %>% 
    group_by(QUD, exh, nonExh)

  transform_to_probs(predictors, simplify=T) %>% add_column(draw = df$.draw)
}

rf = posterior_draws %>% rowwise()
fn <- paste(target_dir, "posterior-samples-probabilities.rds", 
            sep=.Platform$file.sep)
if(file.exists(fn)) { posterior_ps = readRDS(fn)
} else{
  posterior_ps = rf %>% do(p=posterior_draws_probs(.)) %>% 
    unnest(cols = c(p)) %>% group_by(QUD,exh, nonExh, prediction)
  saveRDS(posterior_ps, file = fn)
}

posterior_ps %>% summarize(se=sd(p), p=mean(p))
posterior_samples = posterior_ps %>%
  pivot_wider(names_from = "prediction", values_from = "p")
```

### Main hypotheses H1.1 & H1.2

```{r}
# H1.1 (selection: exhaustive)
h1_exh = posterior_samples %>% dplyr::select(QUD, exh, nonExh, draw, E) %>% 
  pivot_wider(names_from = "QUD", values_from = "E") %>% filter(!is.na(willq)) 
# for each stimulus
h1_exh %>% summarize(p_hypothesis = mean(willq > ifp))
# overall
h1_exh %>% ungroup() %>% summarize(p_hypothesis = mean(willq > ifp))

# H1.2 (selection: both)
h1_both = posterior_samples %>% dplyr::select(QUD, exh, nonExh, draw, both) %>% 
  pivot_wider(names_from = "QUD", values_from = "both") %>% filter(!is.na(willq))
# for each stimulus
h1_both %>% summarize(p_hypothesis = mean(ifp > willq))
# overall
h1_both %>% ungroup() %>% summarize(p_hypothesis = mean(ifp > willq))
```

### Exploratory hypotheses H2 + H3

H2: P(exh | QUD=will-q) > P(both | QUD=will-q)
H3: P(both | QUD=if-p) > P(exh | QUD=if-p)


```{r}
# H2 QUD: will-q
h2 = posterior_samples %>% filter(QUD == "willq")
h2 %>% summarize(p_h2 = mean(E > both)) # for each stimulus                 
h2 %>% ungroup() %>% summarize(p_h2 = mean(E > both)) # across stimuli 

# H3 QUD: if-p
h3 = posterior_samples %>% filter(QUD == "ifp") 
h3 %>% summarize(p_h3 = mean(both >= E)) # for each stimulus
h3 %>% ungroup() %>% summarize(p_h3 = mean(both > E)) # across stimuli
```

Sanity check: P(nonExh) = 0

```{r}
df.exploratory <- posterior_ps %>% 
  mutate(stimulus = case_when(exh == "woD" & nonExh == "int" ~ "A", 
                              exh == "woD" & nonExh == "ext" ~ "B",
                              exh == "withD" & nonExh == "int" ~ "C",
                              exh == "withD" & nonExh == "ext" ~ "D")) 

df.exploratory %>% 
  ggplot(aes(x=p, color = stimulus)) + geom_density() + 
  facet_wrap(QUD~prediction, scales = "free") +
  labs(x = "predicted selection probability")

df.exploratory.ne = df.exploratory %>% filter(prediction == "NE") %>% 
  mutate(is_zero = p <= 0.05) 
  
# stimuli separately
df.exploratory.ne %>% group_by(QUD, stimulus) %>% 
  summarize(p_hypothesis=mean(is_zero)) %>% arrange(desc(p_hypothesis))

# across stimuli
df.exploratory.ne %>% ungroup() %>% summarize(p_hypothesis=mean(is_zero))
```


##  Reaction times

Look at reaction times split across all picture pairs

```{r}
N=100
n_participants = data %>% distinct_at(vars(c(submission_id))) %>% nrow()
critical.RT = data.critical %>% 
  dplyr::select(submission_id, QUD, RT, pic_condition, selected_pic, stimulus) %>%
  mutate(visual_diff = case_when(stimulus %in% c("B", "C") ~ T, 
                                 T ~ F)) %>% 
  group_by(QUD, pic_condition) 
critical.RT_means = critical.RT %>% 
  summarize(mean_RT = mean(RT), .groups = "drop_last")

bootstrap_RTs = group_map(critical.RT, function(df, df.group){
  bootstrap = map(seq(1, N), function(i){
    tibble(mean_RT = mean(sample(df$RT, n_participants, replace=T)))
  }) %>% bind_rows() %>% 
    add_column(pic_condition = df.group$pic_condition, QUD = df.group$QUD) %>% 
    arrange(mean_RT)
  
  N = bootstrap %>% nrow()
  CI.bounds = c(low=ceiling(0.025 * N), up=ceiling(0.975 * N))

  ci.low = bootstrap[CI.bounds[["low"]], ]$mean_RT
  ci.up = bootstrap[CI.bounds[["up"]], ]$mean_RT
  bootstrap %>% add_column(ci.low=ci.low, ci.up=ci.up)
}) %>% bind_rows() %>% group_by(pic_condition, QUD)

bootstrap_RTs.CI.question_pic = bootstrap_RTs %>%
  dplyr::select(-mean_RT) %>% distinct()

left_join(critical.RT_means, bootstrap_RTs.CI.question_pic) %>% 
  ggplot(aes(x=mean_RT, y=pic_condition, fill=QUD)) + 
  geom_bar(stat="identity", position = position_dodge()) + 
  geom_errorbar(aes(xmin=ci.low, xmax=ci.up), position = position_dodge())# +
```

```{r}
p.rt = critical.RT %>%
  ggplot(aes(x=log(RT), color=pic_condition)) + geom_density() +
  guides(color = guide_legend(ncol=2))  +
  scale_color_brewer(name = "stimulus", palette = "Set2")

p.rt
p.rt + facet_wrap(~QUD)
ggsave(file.path(plot_dir, "empiric-rts.pdf"), p.rt, width=7, height=4)
```

```{r}
df.rt = critical.RT %>% 
  dplyr::select(submission_id, QUD, RT, pic_condition, stimulus) %>% 
  #filter(!submission_id %in% ids.scores) %>% 
  separate(col = pic_condition, into = c("exh_situation", "nonExh_situation"), sep="-") %>% 
  mutate(exh_situation = factor(exh_situation, 
                                levels = c("withDistractor", "w/oDistractor")))

df.rt_means = df.rt %>% 
  group_by(QUD, exh_situation, nonExh_situation, stimulus) %>% 
  summarize(mean_rt = mean(RT), .groups = "drop_last")

p <- df.rt_means %>% ggplot(aes(x=exh_situation, y=mean_rt, color=nonExh_situation)) +
  geom_point(aes(shape = QUD)) +
  geom_line(aes(group=interaction(QUD, nonExh_situation))) +
  labs(y = "average reaction time", x = "exhaustive situation") +
  scale_color_brewer(name = "non-exhaustive situation", palette = "Set2")
  # annotate("text", x=rep("withDistractor", 2), y=c(14000, 16000), 
  #               label=c("C", "D"), hjust=2) +
  # annotate("text", x=rep("w/oDistractor", 2), y=c(13500, 15500), 
  #               label=c("A", "B"), hjust=-2)
p 
ggsave(file.path(plot_dir, "reaction-times-233.pdf"), p, width=7, height=4)

```


